---
title: "Quick Start Guide to rlookup"
author: Iqis Gnahz
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

- Use `rlookup` to create a lookup table from a data frame.
- Make desired change to the lookup table file.
- Apply the lookup table to the data frame. 

Attach and load `rlookup`. The `%>%` piping operator from `magrittr` is strongly recommended.
```{r}
library(rlookup)
library(magrittr)
```



## Example data
We are using the good old `mtcars` data to illustrate the use of this package.

```{r}
mtcars <- mtcars %>% tibble::rownames_to_column("model")
head(mtcars)
```



## Make a lookup table

We desire to recode `cyl`, `vs`, `am` variables, which are respectively cylinder count, engine configuration, and transmission type.

A lookup table is called a `lookup` in `rlookup`; it is a data frame that has three columns: `col_name`, `old_value` and `new_value`, containing respectively:

- column name 
- distinct values of the column, to be replaced with
- manually specified values for replacement

Use `build_lookup()` function to select desired columns from a data frame to make a lookup table. The `new_value` column is empty.

```{r}
mtcars_lookup <- mtcars %>% build_lookup(cyl, vs, am)
mtcars_lookup
```



## Write and read lookup table to file

Specify the path in `write_lookup()`. Give file name a `csv` suffix, as the output is in CSV format.


```{r}
# Write the lookup table
# mtcars_lookup %>% write_lookup("mtcars_lookup.csv")

# Examine the file by reading it
# read_lookup("mtcars_lookup.csv")

```

## Edit the lookup table

Edit the `new_value` column using a spreadsheet program, such as OpenOffice Calc or Microsoft Excel; if unavailable, use a free CSV editor such as {Ron's Editor}[https://www.ronsplace.eu/Products/RonsEditor]. 

From your R session, `edit_lookup()` is available to make a quick edit. If the funciton is supplied with a lookup table data frame, it will return the modified lookup table;If supplied with a file path to a lookup table file, modifications will be directly saved to the file. In both instances, the user will be asked to confirm the change.

You may delete value sets (rows) if you don't want to map them.

Do not try anything funny. As with many R programs, `rlookup` is not designed to protect the user from himself.

```{r}
# Let's pretend that you've made desired change to the `new_value`.
read_lookup("mtcars_lookup_edited.csv") %>% write_lookup("mtcars_lookup.csv", overwrite = TRUE)
read_lookup("mtcars_lookup.csv")

```

## Replace values

Now the lookup table is ready, finally it's time to replace values in the the data frame!


```{r}
mtcars_recoded <- mtcars %>% use_lookup("mtcars_lookup.csv")
mtcars_recoded

```

Every value in their corresponding column has changed. That's basically it! 
